Notes on filtration of logics containing {\bf K5}

### 2020-07-06 filtration-stuff/Text.tex
strict filtrations for K5

  \begin{enumerate}
    \item For each $x, y, z \in W$, $x R y$ and $x R z$ implies either $y R z$ or $z R y$.
    \item $R \subseteq R ; R$, that is, $R$ is dense.
    \item For each $x \in W$, $R^{*}(x) = \{ x\} \cup R(R(x))$.
  \end{enumerate}
\end{prop}

@@

If $\Phi$ is finite subset of $\Gamma$ and $\sim = \sim_{\Phi}$, then $\widehat{M}$ is a definable $\Gamma$-filtration of $\mathcal{M}$ through $\Phi$. If $\sim = \sim_{\Gamma}$, then such a filtration by means of the definiton above is called \emph{strict}.

\begin{lemma}
  Let $\Gamma$ be a finite set of formulas closed under subformulas and $\widehat{M}$ a filtration of $\mathcal{M}$ through $\Gamma$, then for each $x \in W$ and for each $\varphi \in \Gamma$ one has
  \begin{center}
    $\mathcal{M}, x \models \varphi \Leftrightarrow \widehat{M}, \hat{x} \models \varphi$
  \end{center}
\end{lemma}

@@

@@

>>>>

  \begin{enumerate}
    \item For each $x, y, z \in W$, $x R y$ and $x R z$ implies either $y R z$ or $z R y$.
    \item $R \subseteq R ; R$, that is, $R$ is dense.
    \item For each $x \in W$, $R^{*}(x) = \{ x\} \cup R(R(x))$.
    \item $R^{-1} ; R \subseteq R$.
  \end{enumerate}
\end{prop}

@@

If $\Phi$ is finite subset of $\Gamma$ and $\sim = \sim_{\Phi}$, then $\widehat{\mathcal{M}}$ is a definable $\Gamma$-filtration of $\mathcal{M}$ through $\Phi$. If $\sim = \sim_{\Gamma}$, then such a filtration by means of the definiton above is called \emph{strict}.

\begin{lemma}
  Let $\Gamma$ be a finite set of formulas closed under subformulas and $\widehat{\mathcal{M}}$ a filtration of $\mathcal{M}$ through $\Gamma$, then for each $x \in W$ and for each $\varphi \in \Gamma$ one has
  \begin{center}
    $\mathcal{M}, x \models \varphi \Leftrightarrow \widehat{\mathcal{M}}, \hat{x} \models \varphi$
  \end{center}
\end{lemma}

@@

\subsection{Strict filtration for logics containing ${\bf K5}$}

\begin{defin}
  Let $\mathcal{F} = \langle W, R, \vartheta \rangle$ be a Kripke model. Then its Euclidean Horn closure ($R^E$, $E$-closure) is defined as
  \begin{center}
    $R^E = (\bigcup \limits_{n < \omega} (R^{-1})^n); R$
  \end{center}
  or, equivalently, $R^{E} = \bigcup \limits_{n < \omega} R_n$, where $R^{0} = R$ and $R_{n + 1} = R_n \cup (R^{-1} ; R_n)$.
\end{defin}

\begin{lemma}
  Let $\mathcal{M} = \langle W, R, \vartheta \rangle$ be an Euclidean model, $\Gamma$ a set of $\operatorname{Sub}$-closed formulas, and $\sim$ an equivalence relation having a finite index that respects $\Gamma$. Then its $E$-closure of the minimal filtration of $R$ is a filtration itself.
\end{lemma}

\begin{proof}
  Let us show that $\widehat{R^{E}} = \bigcup \limits_{n < \omega} (R_n)^{\text{min}}_{\sim}$ is a filtration and $\langle W / \sim, \widehat{R}^{E}, \widehat{\vartheta} \rangle$ is an Euclidean model.

  Let $\widehat{\mathcal{M}} = \langle \widehat{W}, \widehat{R}, \widehat{\vartheta} \rangle$ be a minimal filtration of an Euclidean model through $\sim$.

  \begin{enumerate}
    \item Suppose $x R y$, then $[x] \widehat{R} [y]$ by the definition of a minimal filtration. $\widehat{R}$ is clearly a subrelation of $\widehat{R}^{E}$, thus $[x] \widehat{R}^{E} [y]$.
    \item Suppose $[x] \widehat{R^{E}} [y]$ and $\mathcal{M}, x \models \Box \varphi$ for $\Box \varphi \in \Gamma$. then $([x], [y]) \in \bigcup \limits_{n < \omega} R_n$. Then $([x], [y]) \in R_n$ for some $n < \omega$.
    There are two cases:
    \begin{enumerate}
      \item $n = 0$, then $[x] \widehat{R} [y]$, so, obviously, one has $\mathcal{M}, y \models \varphi$
      \item $n = m + 1$. Suppose $[x] {\widehat{R}}_{m + 1} [y]$. Thus, $[x] \widehat{R}_n \cup (\widehat{R}^{-1} ; \widehat{R}_n) [y]$.
      There are the following two cases:
      \begin{enumerate}
        \item $([x], [y]) \in \widehat{R}_n$, then the statement holds by IH.
        \item $([x], [y]) \in \widehat{R}^{-1} ; \widehat{R}_n$.
        Then there exists $[z]$ such that $[x] \widehat{R}^{-1} [z]$ and $[z] \widehat{R}_m [y]$, that is,
        $[z] \widehat{R} [x]$ and $[z] \widehat{R}_m [y]$. One has $\mathcal{M}, x \models \Box \varphi$, then
        $\widehat{\mathcal{M}}, [x] \models \Box \varphi$. So $\widehat{\mathcal{M}}, [z] \models \Diamond \Box \varphi$. On the other hand, $\mathcal{M}, z \models \Diamond \Box \varphi \to \Box \varphi$, so $\mathcal{M}, z \models \Box \varphi$ and,
        from that, $\widehat{\mathcal{M}}, [z] \models \Box \varphi$, and, thus, $\widehat{\mathcal{M}}, [y] \models \varphi$ and $\mathcal{M}, y \models \varphi$ by IH and the definition of a minimal filtration.
      \end{enumerate}
    \end{enumerate}
  \end{enumerate}
\end{proof}

\begin{col}
  ${\bf K}5$ admit strict filtrations.
\end{col}

@@

\subsection{}
